\setuppapersize[letter][letter]

\setuplayout[backspace=1.5in,width=5.5in,margin=2.5em]

\setuppagenumbering[location=footer]

\setupinmargin[align=right]

\setupitemize[1][paragraph,atmargin,unpacked][symbol=n,indentnext=no,]
\setupitemize[2][symbol=a][indentnext=no]
%\setupitemize[3][symbol=a][indentnext=no]

\definehead[CVHEAD][subject]
\setuphead[subject][style=\bfa,after={\blank[medium]}]

\definehead[SUBCVHEAD][subsubject]
\setuphead[subsubject][style=\bf,after={\blank[small]},before={\blank[small]}]

\defineparagraphs[ContactColumns][n=2]
\setupparagraphs[ContactColumns][each][rule=off,distance=0pt]

\define[2]\CVTITLE{\midaligned{\framed[frame=off,align=middle]
                                {{\bfa  
#1}\blank[medium]{\tfb#2}}}\blank[small]}

\define[1]\CVEMAIL{{\bf Email:}{ \type{#1}}\blank}

% \framed is good for testing or fine-tuning \CONTACT
\define[2]\CONTACT{%\framed[frame=off,width=broad,align=right,strut=no,
                     %offset=none,height=fit]
                     %{%
                    \startalignment[middle]
                    \startContactColumns
                     {\bf Office:}\crlf
                     {#1}
                     \ContactColumns
                     {\bf Home:}\crlf
                     {#2}
                    \stopContactColumns
                    \stopalignment}
                    %}

% \framed is used to prevent pagebreaks within an \<>ITEM<>

\define[5]\EDUITEM{\framed[frame=off,align=right,width=broad]
                     {{\bf #1}\inmargin{\it #2}\ (#3).\crlf
                     {\bf #4}, #5.\blank[small]}}
\define[6]\EDUITEMMA{\framed[frame=off,align=right,width=broad]
                     {{\bf #1}\inmargin{\it #2}\ (#3).\crlf
                     {\bf #4}, #5.\crlf

\starttext

\CVTITLE{Curriculum Vitae}{Idris Samawi Hamid}

\CONTACT{Department of \TeX\ Studies\crlf
          % 221 Eddy Hall\crlf
          \TeX\ State University\crlf
          Fort Collins, CO 80523 \crlf USA\crlf
          (970) 491-2406/6315
         }
         {1234 \TeX\ Road\crlf
          \TeX{}ville, CO 80555 \crlf USA\crlf
          (800) TEX-5555\crlf}

\CVEMAIL{ishamid  texstate.edu}

\CVHEAD{Education}

\EDUITEMPHD
     {PhD}
     {2007}
     {\TeX\ Studies}
     {\TeX\ University}
     {Amherst, NY}
     {Arabic Critical Editions}
     {\TeX\ Distinguished Professor Hans Hagen}

\EDUITEMMA
     {MA}
     {1996}
     {Physics}
     {University at Buffalo, State University of New York}
     {Amherst, NY}
     {G\"{o}del's Universe, the Mach Principle, and the Cosmological  
Principle:
     Considerations for the Philosophy of Physics}

\EDUITEM{BSc}{1990}{Physics; minor, Mathematics}{Georgia State
University}{Atlanta, GA}

%% positions
% academic positions
\CVHEAD{Academic Positions}

\ACADITEM
     {2001--2007}
     {\TeX\ State University, \TeX{}ville, CO}
     {Associate Professor}
     {Instruction and research, especially
     the development of upper-level courses in \TeX\ and \Context}

\CVHEAD{Areas of Specialization}

Metaphysics of \TeX, Islamic Philosophy, Philosophy of Mysticism,
Philosophy of Science.

\CVHEAD{Philosophic Interests}

\TeX\ Logic, Category Theory, Chinese Philosophy.

% other skills

\CVHEAD{Languages and Technical Skills}

Arabic, French, and the professional computer typesetting language
\TeX, including the major dialects \LaTeX, \Context, as well as
the minor dialects Arab\TeX (for typesetting in Arabic script) and
EDMAC (for critical edition typesetting). I am also developing a
new Arabic-script system for scholarly typesetting, {\sc Oriental
\TeX}, based upon \TeX\ and \Context.

% patents

\CVHEAD{Publications}

\SUBCVHEAD{Books Authored}

\startitemize

\item Idris Samawi Hamid, 2006, {\it The Cosmology of \TeX}, \TeX\
University Press, 240 pp.

\stopitemize

\SUBCVHEAD{Articles}

\startitemize

\item Idris Samawi Hamid,  2005, {\it Installing Expert Fonts:
Minion Pro}. In {\it MAPS}, Issue~33; pp.~19--35. Published by the
Nederlandstalige \TEX\ Gebruikersgroep (NTG), Netherlands.

\stopitemize

\CVHEAD{Grants}

\GRANTITEM{2006--2007} {{\it Oriental \TeX: A New Direction in
Scholarly Complex-Script Typesetting}; Academic Enrichment
Program, College of Liberal Arts, CSU; \$42,927}

\CVHEAD{Presentations}

\startitemize
     \item 2007, ``Oriental \TeX'', 1\high{st} Annual \Context\ Conference,
     Epen, the Netherlands.

\stopitemize

\CVHEAD{Service}

\SUBCVHEAD{Profession}

\startitemize

\item The \TeX\ Users Group.

\stopitemize

\SUBCVHEAD{Department}

\startitemize
     \item Graduate Committee,  2002--2005, 2006--2007.
     \item Advisory Committee,  2005--2006.
\stopitemize

\SUBCVHEAD{College}

\startitemize
     \item \TeX\ Studies Committee, 2001--2006.
\stopitemize

\SUBCVHEAD{University}

\startitemize
     \item Faculty Council, Fall 2005.
\stopitemize

\CVHEAD{Thesis/Dissertation Direction}

\startitemize
     \item  Joseph Ramagli (MA), ``\TeX\ Theology in Context'',
     in progress.
\stopitemize

\stoptext
